---
title: "README" 
output:  
github_document: 
pandoc_args: ["--wrap=none"] 
always_allow_html: true
---


### **Dataset Overview**  

**<strong>When was the data collected?</strong>**  

The data was collected in **2014**, with a time range from **May 2, 2014, to July 10, 2014**. It was uploaded to Kaggle and last updated **seven months ago**.  

---

**<strong>Where was the data acquired?</strong>**  

The data was acquired from **[Kaggle](https://www.kaggle.com)**. The original source was **[Zillow](https://www.zillow.com)**, a real estate website known for its housing market data.  

---

**<strong>How was the data acquired?</strong>**  

Zillow’s **Economic Research Team** gathers, refines, and publishes housing and economic data from both **public and proprietary sources**. The core of Zillow’s data comes from:  

- **Public property records** (deeds, parcel information, transaction history).  
- **Internal Zillow market analyses** using proprietary algorithms.  
- **Government and private housing reports** for contextual insights.  

These sources are used to calculate various housing metrics, explained in the next section.  

---

### **Dataset Attributes**  

**<strong>What are the attributes of this dataset?</strong>**  

The dataset consists of **18 attributes** that describe various characteristics of properties:  

- **Date:** The date the property was sold.  
- **Price:** The sale price of the property in USD, serving as the target variable in housing market analyses.  
- **Bedroom:** The total number of bedrooms in the property, indicating the home’s capacity.  
- **Bathroom:** The total number of bathrooms in the property, including full and half-baths.  
- **Sqft_living:** The total interior square footage of the home, representing the livable space.  
- **Sqft_lot:** The total land area of the property, including the house and yard.  
- **Floors:** The number of floors in the home, influencing the layout and design.  
- **Waterfront:** A binary variable indicating whether the property is located on the waterfront (1 = Yes, 0 = No).  
- **View:** An index ranging from **0 to 4**, where higher values indicate a better quality view from the property.  
- **Condition:** An index ranging from **1 to 5**, where **1** represents poor condition and **5** represents excellent condition.  
- **Sqft Above:** The total square footage of the home **excluding** the basement, reflecting the main living area.  
- **Sqft Basement:** The total square footage of the basement area, which may or may not be finished living space.  
- **Yr Built:** The year in which the property was originally constructed.  
- **Yr Renovated:** The most recent year in which the property underwent major renovations or updates.  
- **Street:** The street address of the property.  
- **City:** The city where the property is located.  
- **Statezip:** A combined variable containing both the **state** and **zip code** of the property.  
- **Country:** The country where the property is located.  

---

### **Data Types**  

**<strong>What type of data do these attributes contain?</strong>**  

| **Data Type**  | **Attributes** |
|---------------|----------------|
| **Nominal**   | Street, City, Statezip, Country, Waterfront |
| **Ordinal**   | View, Condition |
| **Interval**  | Yr Built, Yr Renovated |
| **Ratio**     | Price, Bedroom, Bathroom, Sqft_living, Sqft_lot, Floors, Sqft Above, Sqft Basement |

- **Nominal:** Categorical variables without a meaningful order, such as property location attributes.  
- **Ordinal:** Ranked variables with a meaningful order but uneven intervals, such as **View** and **Condition** ratings.  
- **Interval:** Numeric variables with meaningful differences but no true zero, such as **Yr Built** and **Yr Renovated**.  
- **Ratio:** Continuous numerical attributes with a true zero, such as **Price**, **Sqft_living**, and **Sqft_lot**.  

```{r}
library(knitr)
library(kableExtra)

# Create a table for dataset attributes
attributes_table <- data.frame(
  Attribute = c("Date", "Price", "Bedroom", "Bathroom", "Sqft_living", "Sqft_lot", "Floors", "Waterfront",
                "View", "Condition", "Sqft Above", "Sqft Basement", "Yr Built", "Yr Renovated", "Street", "City",
                "Statezip", "Country"),
  Description = c("Date the property was sold",
                  "Sale price of the property in USD (target variable)",
                  "Number of bedrooms in the property",
                  "Number of bathrooms in the property (full and half-baths)",
                  "Total interior square footage of the home",
                  "Total land area of the property, including the house and yard",
                  "Number of floors in the home",
                  "Binary indicator for waterfront location (1 = Yes, 0 = No)",
                  "Index (0-4) rating the quality of the property's view",
                  "Index (1-5) rating the condition of the property",
                  "Total square footage of the home excluding the basement",
                  "Total square footage of the basement area",
                  "Year the property was originally constructed",
                  "Year of the most recent renovation",
                  "Street address of the property",
                  "City where the property is located",
                  "State and ZIP code of the property",
                  "Country where the property is located"),
  DataType = c("Date", "Ratio", "Ratio", "Ratio", "Ratio", "Ratio", "Ratio", "Nominal",
               "Ordinal", "Ordinal", "Ratio", "Ratio", "Interval", "Interval", "Nominal", "Nominal",
               "Nominal", "Nominal")
)

# Format and display the table
kable(attributes_table, caption = "Dataset Attributes and Descriptions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "15%") %>%
  column_spec(2, italic = TRUE, width = "65%") %>%
  column_spec(3, width = "20%") %>%
  row_spec(0, background = "lightgray")
```

```{r}
# Load necessary libraries
library(dplyr)
library(readr)
library(knitr)
library(kableExtra)

# Load the dataset (make sure the CSV is in your working directory or adjust the path)
df <- read_csv("USA Housing Dataset.csv")

# Select numeric columns
numeric_df <- df %>%
  select(where(is.numeric))

# Generate summary statistics
numeric_summary <- data.frame(
  Variable = names(numeric_df),
  Count = sapply(numeric_df, function(x) sum(!is.na(x))),
  Mean = sapply(numeric_df, mean, na.rm = TRUE),
  SD = sapply(numeric_df, sd, na.rm = TRUE),
  Min = sapply(numeric_df, min, na.rm = TRUE),
  Q1 = sapply(numeric_df, quantile, probs = 0.25, na.rm = TRUE),
  Median = sapply(numeric_df, median, na.rm = TRUE),
  Q3 = sapply(numeric_df, quantile, probs = 0.75, na.rm = TRUE),
  Max = sapply(numeric_df, max, na.rm = TRUE)
)

# Display as a formatted table using kable
kable(numeric_summary, digits = 2, caption = "Summary Statistics - Numeric Variables") %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(1, bold = TRUE)
```



